### 2025-03-22 latex/paper.tex
tex init

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{geometry}
\geometry{legalpaper, portrait, margin=.5in}
\title{CS 2051 Group Project}
\author{}
\date{March 2025}

\begin{document}

\maketitle
\end{document}